Looking back at the motivation, the goal of outperforming QEMU~\cite{bellard2005qemu} is achieved consistently by \translatorname{}.
The performance is measured using \textit{SPEC CPU 2017}~\cite{spec-cpu-2017} an industry-standardized, CPU intensive package of benchmark suites for measuring and comparing compute-intensive performance.
In most cases we can achieve a execution time of around twice the native, but the tests also show areas with further optimisation possibilities. %todo speed seems strange here makes it sound like we are faster than native @Noah
The quite significant performance increases in comparison to QEMU can be attributed to the specialisation in RISC--V to x86--64 translation leading to a straight forward design of the translation process and enabling specific optimisations for this guest-host pairing.

The benchmarks suites did not only help to ensure the performance of \translatorname{}, but equally important they tested the implementation in various scenarios and helped to discover bugs due to their assertion of the test programs' outputs.
Furthermore many unit tests were designed to assure the correct translation of (almost) every RISC--V instruction in diverse contexts.
All in all nearly 30.000 test cases are being executed by our own test suites based on~\textit{Google Test}~\cite{gtest} covering the RISC--V parser as well as the emitted x86--64 assembly.

The development of \translatorname{} focused on supporting and optimizing the core instruction set and integer extensions so far.
Support for floating point extensions however is already implemented and tested.
It will be benchmarked by~\textit{SPEC CPU 2017} as soon as possible.
Our preliminary tests nonetheless already show a huge performance increase in comparison to QEMU as a consequence of QEMU's emulation and \translatorname{}'s use of the native SSE--extension for floating point arithmetic.


Currently \translatorname{} supports all base instruction sets and extensions needed for executing single threaded programs besides the \textit{"Q" Standard Extension}.
Future standard extensions such as the \textit{"B" Standard Extension} for bit manipulation can easily be added to \translatorname{} when they come available by adding decoding of this instructions to the parser and writing tailored translator functions.
Support of multithreading would be a bigger undertaking as it requires a redesign of \translatorname{}'s core to support the necessary memory consistency and atomicity of some memory access (\textit{"A" Standard Extension}).
Thus, this is not planned to be added in the near future.

Furthermore some operating system related features are not provided yet.
Specifically only a subset of the existing system calls is implemented and loading of dynamically linked libraries at runtime is not supported.

The next steps for generally improving \translatorname{} would be implementing techniques used for integer arithmetic (e.g.\ lazy register replacement) for floating point arithmetic as well and supporting even more system calls.
Besides this one could invest into one of the aforementioned areas specifically, for example applying a more rigid memory consistency model.

Concerning the performance optimisation most development time has been invested on optimising the emitted code and reduction of context switches.
Benchmarking and optimization of the parsing and translation time has mostly been left out so far and might be due to optimization.


The current version of \translatorname{} is capable of outperforming QEMU whilst providing almost the same functionality.
In some aspects, e.g.\ floating-point arithmetic, we are even ahead of QEMU by using native instructions from the SSE-extension instead of relying on emulation.
Also there is still potential for improvement the \textit{SPEC CPU 2017} benchmarks demonstrates its usability in executing general purpose RISC--V compiled programs on a x86--64 processor. %TODO what? @Noah